Supporting Information for "Ship-based lidar evaluation of Southern Ocean low clouds in the storm-resolving general circulation model ICON, and the ERA5 and MERRA-2 reanalyses"

### 2025-07-18 manuscript_agu_rev1.tex
Various

\subsection{Top of Atmosphere Radiation, Liquid and Ice Water Path}

@@

Fig.~\ref{fig:stats-hist} shows the SW and LW radiation, liquid water path (LWP), and ice water path (IWP) distributions as histograms and their corresponding averages. LWP and IWP are calculated from the mass of water in the column divided by the area of the column, i.e. not just the area of the cloudy portion of the column, as in some definitions. ERA5 and ICON overestimate outgoing SW near 80 Wm$^{-2}$ (Fig.~\ref{fig:stats-hist}a), which probably relates to clear sky situations, as expected from the underestimated cloud fraction. They also underestimate the highly reflective situations above 200 Wm$^{-2}$. MERRA-2 exhibits the too few, too bright problem in terms of overestimating SW reflectivity around 290 Wm$^{-2}$, given that the total cloud fraction in MERRA-2 is strongly underestimated. The LW distribution shows that all of the models overestimate outgoing LW (Fig.~\ref{fig:stats-hist}b), which is expected from the underestimated cloud fraction, exposing more of the warmer ocean surface relative to colder clouds. The LWP distribution shows that all models overestimate cases with near-zero LWP (Fig.~\ref{fig:stats-hist}c), which relates to the underestimated total cloud fraction. MERRA-2 shows quite overestimated high-LWP situations, which is most likely related to the too few, too bright problem of simulating lower total cloud fraction but clouds with higher LWP to compensate. IWP (Fig.~\ref{fig:stats-hist}d) is somewhat less important radiatively than LWP because of the typically larger and less numerous hydrometeors. Similarly to LWP, the models overestimate situations with near-zero IWP. ERA5 is otherwise simulating the IWP distribution well, but ICON and MERRA-2 underestimate IWP. Fig.~\ref{fig:stats-hist-cloudy} shows the same as Fig.~\ref{fig:stats-hist}, but for cloudy profiles only. This figure shows more distinctly that in cloudy situations, MERRA-2 overestimates moderate (0.05--0.15 kg m$^{-2}$) and high LWP (over 0.15 kg m$^{-2}$), and ERA5 and ICON underestimate moderate LWP. ICON also overestimates high LWP, resulting in overestimated average LWP.

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth]{img/stats_hist.pdf}
\caption{
Histograms and averages of outgoing \textbf{(a)} SW and \textbf{(b)} LW radiation at TOA, \textbf{(c)} liquid water path, and \textbf{(d)} ice water path in CERES SYN1deg observations (OBS), ICON, MERRA-2, and ERA5 along the voyage tracks. All voyages are weighted equally. The statistics are calculated from daily mean values corresponding to each time step and geographical location of the voyage tracks.
}
\label{fig:stats-hist}
\end{figure}

@@

The radiosonde observations indicate that the LCL is too high in ICON and reanalyses, which is probably responsible for the higher peak of clouds in the models and the lack of very low-level clouds and fog. Notably, ICON exhibits smaller internal variability in $\theta_v$ than the radiosonde observations. The analysis of cloud liquid and ice water path (Fig.~\ref{fig:stats-hist} and \ref{fig:stats-hist-cloudy}) shows that both phases are present in observations in about equal amounts. The models show diverse biases, the most pronounced being overestimation of high-LWP values in MERRA-2 and overestimation of cases with near-zero LWP and IWP in all models. All models tend to compensate for the overestimated cases of near-zero LWP with more high-LWP values to get a mean LWP that is either less (but close) to the observations (ERA5) or higher than the observations (ICON and MERRA-2). IWP is underestimated in all of the models. In the case of ICON and MERRA-2, the mean IWP was underestimated and LWP overestimated, indicating that the models produce too much liquid and not enough ice phase. This is in contrast with previous findings of the lack of supercooled liquid over the SO in other models. If the liquid phase is overestimated relative to the ice phase, one would expect underestimated cloud SW reflectivity due to a larger number of smaller hydrometeors for the same amount of water. Cloudy areas would then appear brighter in the SW spectrum. This can contribute to the too few, too bright bias, i.e., the the overestimated brightness of cloudy areas compensates for the lower total cloud fraction in the models.

>>>>

\subsection{Top of Atmosphere Radiation, Liquid and Ice Water Path}
\label{sec:rad-lwp-iwp}

@@

Fig.~\ref{fig:stats-hist-rad} shows the SW and LW radiation as histograms and their corresponding averages. ERA5 and ICON overestimate outgoing SW near 80 Wm$^{-2}$ (Fig.~\ref{fig:stats-hist-rad}a), which probably relates to clear sky situations, as expected from the underestimated cloud fraction. They also underestimate the highly reflective situations above 200 Wm$^{-2}$. MERRA-2 exhibits the too few, too bright problem in terms of overestimating SW reflectivity around 290 Wm$^{-2}$, given that the total cloud fraction in MERRA-2 is strongly underestimated. The LW distribution shows that all of the models overestimate outgoing LW (Fig.~\ref{fig:stats-hist-rad}b), which is expected from the underestimated cloud fraction, exposing more of the warmer ocean surface relative to colder clouds.

Fig.~\ref{fig:stats-hist-si} shows liquid water path (LWP), and ice water path (IWP) distributions as histograms and their corresponding averages. LWP and IWP are calculated from the mass of water in the column divided by the area of the column, i.e. not just the area of the cloudy portion of the column, as in some definitions. The available observational satellite reference for LWP and IWP over high latitudes is unfortunately very uncertain due to a high solar zenith angle and inability of passive visible and infrared retrievals to detect phase below the cloud top of mixed-phase clouds \cite{huang2006,greenwald2009,seethala2010,eliasson2011,duncan2018,khanal2020}, and this limits our comparison. The LWP distribution shows that all models overestimate cases with near-zero LWP (Fig.~\ref{fig:stats-hist-si}a), which relates to the underestimated total cloud fraction. MERRA-2 shows quite overestimated high-LWP situations, which is most likely related to the too few, too bright problem of simulating lower total cloud fraction but clouds with higher LWP to compensate. IWP (Fig.~\ref{fig:stats-hist-si}b) is somewhat less important radiatively than LWP because of the typically larger and less numerous hydrometeors. Similarly to LWP, the models overestimate situations with near-zero IWP. ERA5 is otherwise simulating the IWP distribution well, but ICON and MERRA-2 underestimate IWP. In the cloudy situations (Fig.~\ref{fig:stats-hist-si}c, d), it can be seen more distinctly that MERRA-2 overestimates moderate (0.05--0.15 kg m$^{-2}$) and high LWP (over 0.15 kg m$^{-2}$), and ERA5 and ICON underestimate moderate LWP. ICON also overestimates high LWP, resulting in overestimated average LWP.

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth]{img/stats_hist_rad.pdf}
\caption{
Histograms and averages of outgoing \textbf{(a)} SW and \textbf{(b)} LW radiation at TOA in CERES SYN1deg observations (OBS), ICON, MERRA-2, and ERA5. All campaigns are weighted equally. The statistics are calculated from daily mean values corresponding to each time step and geographical location of the voyage tracks and stations.
}
\label{fig:stats-hist-rad}
\end{figure}

@@

The radiosonde observations indicate that the LCL is too high in ICON and reanalyses, which is probably responsible for the higher peak of clouds in the models and the lack of very low-level clouds and fog. Notably, ICON exhibits smaller internal variability in $\theta_v$ than the radiosonde observations. The analysis of LWP and IWP (Fig.~\ref{fig:stats-hist-si}) shows that both phases are present in observations in about equal amounts. The models show diverse biases, the most pronounced being overestimation of high-LWP values in MERRA-2 and overestimation of cases with near-zero LWP and IWP in all models. All models tend to compensate for the overestimated cases of near-zero LWP with more high-LWP values to get a mean LWP that is either less (but close) to the observations (ERA5) or higher than the observations (ICON and MERRA-2). IWP is underestimated in all of the models. In the case of ICON and MERRA-2, the mean IWP was underestimated and LWP overestimated, indicating that the models produce too much liquid and not enough ice phase. This is in contrast with previous findings of the lack of supercooled liquid over the SO in other models. If the liquid phase is overestimated relative to the ice phase, one would expect underestimated cloud SW reflectivity due to a larger number of smaller hydrometeors for the same amount of water. Cloudy areas would then appear brighter in the SW spectrum. This can contribute to the too few, too bright bias, i.e., the the overestimated brightness of cloudy areas compensates for the lower total cloud fraction in the models. As mentioned in Section~\ref{sec:rad-lwp-iwp}, the LWP and IWP are, however, affected by high uncertainty of the satellite retrievals.

### 2025-07-18 si_rev1.tex
Various

\begin{figure}[p!]
\includegraphics[width=\textwidth]{img/stats_hist_cloudy.pdf}
\caption{
The same as Fig.~\ref{fig:stats-hist} but for cloudy profiles only, as identified by the lidar cloud mask.
}
\label{fig:stats-hist-cloudy}
\end{figure}

>>>>

\begin{figure}[p!]
\includegraphics[width=\textwidth]{img/stats_hist_si.pdf}
\caption{
Histograms and averages of \textbf{(a, c)} liquid water path, and \textbf{(b, d)} ice water path in CERES SYN1deg observations (OBS), ICON,
MERRA-2, and ERA5 for \textbf{(a, b)} all and \textbf{(c, d)} cloudy profiles. All campaigns are weighted equally. The statistics are calculated from daily mean values corresponding to each time step and geographical location of the voyage tracks and stations.
}
\label{fig:stats-hist-si}
\end{figure}
